Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2017-11-28 Ben_Manuscripts/structure_paper/Outline.tex
Addressed what comments I could and did some minor reorganization

  While RO and NF have seen many advances in the past few decades, they 
  are far from perfect separation technologies.
  \begin{itemize}
    \item \textit{RO membranes}
    \item Inconsistent performance : Current state-of-the-art RO membranes are unstructured with
    tortuous and polydisperse diffusion pathways which leads to 
    inconsistent performance \cite{song_nano_2011}
    \item High energy requirements : Necessarily high feed pressures 
    drive up energy requirements which strains developing regions and
    contributes strongly to CO\textsubscript{2} emissions. \cite{mcginnis_global_2008}
    \item Separation based on differences in solubility and diffusivity:
%MRS2: couldn't you argue that the fact that they diffuse and dissolve at different rates means that it makes it possible to  design targeted separations?
    Moreover, designing RO membranes to achieve targeted separations of 
    specific solutes is nearly impossible because various solutes dissolve
    into and diffuse through the polymer matrix at different rates. \cite{wijmans_solution-diffusion_1995}
    \item At best, one can exploit these differences to create a functional
    selective barrier.
    \item \textit{NF membranes}
    \item NF was introduced as an intermediate between RO and ultrafiltration,
    having the ability to separate organic matter and salts on the order of 
    one nanometer in size.
%MRS2: is ``well-defined'' the right word if there is a ``pore size distribution''?
    \item Larger and well-defined pores drive down energy requirements while
    still affording separation of solutes as small as ions to some degree \cite{van_der_bruggen_review_2003}
    \item NF is often used as a precursor to reverse osmosis
    \item Unfortunately, NF membranes, like RO, possess a pore size 
    distribution which limits their ability to perform precise separations \cite{bowen_modelling_2002}
  \end{itemize}

  Nanostructured membranes can bypass many of the performance issues which
  plague traditional NF and RO membranes.
  \begin{itemize}
    \item Tune size and functionality of building blocks to control pore
    size and shape: One can accomplish targeted separations with high 
    selectivity by tuning shape, size and functionality of the molecular
    building blocks which form these materials. % BJC: "these materials" --> "nanostructured membranes", or is that redundant? %MRS2: seems fine for now, we can tweak this later. 
    \item As a result, solute rejecting pores can have their sizes tuned
    uniformly, resulting in strict size cut-offs.
    \item Entirely different mechanisms may govern transport in a given
    nanostructured material which can inspire novel separation techniques.
  \end{itemize}

@@

  Self assembling lyotropic liquid crystals (LLCs) are a suitable candidate for
  aqueous separation applications. 
  \begin{itemize}
    \item LLCs share the characteristic ability of nanostructured membrane
    materials to create highly ordered structures with the added benefits
    of low cost and synthetic techniques feasible for 
    large scale production \cite{feng_scalable_2014}
    \item Neat liquid crystal monomer forms the thermotropic, Col\textsubscript{h}
    phase. The presence of water results in the H\textsubscript{II} phase.
    \item In both cases, monomers assemble into mesophases made of hexagonally
    packed, uniform size, cylinders with hydrophilic groups oriented inward
    towards the pore center and hydrophobic groups facing outward.
    \item H\textsubscript{II} and Col\textsubscript{h} phase systems created by
    the monomer named Na-GA3C11 have been extensively studied experimentally \cite{smith_ordered_1997, %BJC: IUPAC chemical name here?
    zhou_supported_2005,resel_h2-phase_2000,feng_scalable_2014,feng_thin_2016}. 
    % BJC: New paragraph?
    %MRS2: shorter paragraphs generally good.
    \item Until recently, mesophases formed by Na-GA3C11 could not be macroscopically
    aligned, resulting in low flux membranes, slowing research in the field.
%MRS2: slowed research because it wasn't clear there was a route to scaleable filtration.
    \item In 2014, Feng et al. showed that the mesophases could be aligned 
    using a magnetic field with subsequent crosslinking to lock the structure
    in place \cite{feng_scalable_2014}
    \item In 2016, Feng et al. showed that the same result could be obtained 
    using a technique termed soft confinement \cite{feng_thin_2016}.
    \item Following this breakthrough, research into LLC membranes has been
    reinvigorated.
%MRS2: not THAT many people are doing it, so hard to say reinvigorated quite yet. Pick other verb.
  \end{itemize}

@@

  A molecular level understanding of LLC membrane structure, enabled by molecular
  dynamics simulations, will provide guidelines to reduce the large chemical space
  available to design monomers for creation of separation-specific membranes.
  \begin{itemize}
%     \item LLCs are versatile and controllable with a \textbf{large chemical design
%     space} available for membrane design  %BJC: redundant
    \item Over the past 20 years, H\textsubscript{II} phase LLC membrane studies have been limited primarily
    to Na-GA3C11 with some characterization done after minor structural modifications
%MRs: specify what type of modification.
    \cite{resel_structural_2000}.
    \item Rejection studies show that this membrane can not perform separations of
    solutes less than 1.2 nm in diameter because the pores are too large \cite{zhou_supported_2005}.
    \item  We do not yet understand how to reduce the effective pore size 
%MRS2: controllably
      or 
    how to tune the chemical environment in the nanopores for effective water
    desalination or small organic molecule separations.
    \item It will be challenging to efficiently narrow down the large design space in 
    a laboratory setting without a robust model.
    \item The only source of predictive modeling for LLC systems have been macroscopic models 
    which likely do not adequately describe transport at these length scales. % BJC: Reference. I think it is just w.r.t. bicontinuous cubic
%MRS2: not quite sure of the phrasing here on next two points. Something more a long the lines of the choice of head group 
may allow tuning rejection of solutes based on size?
    \item Choice of head group may play a role in the rejection of charged or uncharged solutes.
    \item Choice of counterion may influence the establishment of a Donnan potential
    affecting the degree to which the membrane can exclude charged species.
    \item A good molecular model should incorporate a detailed picture of the nanoscopic pore 
    structure which will be crucial to understanding the role of monomer structure in 
    membrane design.
    \item Molecular dynamics simulations will provide the required level of detail
  \end{itemize}

  %BJC: Figure here showing previous level of detail compared to atomistic level of detail
  %MRS2: provenance of the atomistic figure could be explained better in the caption. It also looks disordered. Will eventually need a higher resolution figure for the previous understanding pore.
  \begin{figure}
  \centering
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{nanopore_undetailed.jpg}
		\caption{}~\label{fig:undetailed_pore}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{detailed_pore.png}
		\caption{}~\label{fig:detailed_pore}
	\end{subfigure}
  \caption{(a) Our previous understanding of the pore structure allows us to speculate
	   about separation behavior. (b) A detailed molecular model will allow us to
	   directly observe solute transport}~\label{fig:detail}
  \end{figure}

  Our approach to constructing a general model will follow the development of a
  model of a specific LLC membrane with sufficient experimental characterization.
    \begin{itemize}
	    \item We have chosen to focus on assemblies formed by Na-GA3C11
	    \item We have also narrowed our scope to the development of 
	    a model of the Col\textsubscript{h} phase membrane.
	    \item Compared to the H\textsubscript{II} phase, the Col\textsubscript{h}
	    phase is a simpler starting point, due to the absence of water, and has
	    more detailed experimental structural data.  % BJC: More detailed because of WAXS. HII doesn't have any (that I know of) WAXS
%MRS2: explicitly state the presence of WAXS data.
	    \item Despite having structural data, there is still information which 
	    experiment cannot definitively answer.
            %MRS2: such as?
    \end{itemize}

@@

   \item Is it necessary to include any water in order to appropriately model the 
   Col\textsubscript{h} phase? \label{point:water}
   \begin{itemize}
	\item While the Col\textsubscript{h} phase is described as dry, it is 
	possible that small amounts of ambient water may be leached into the system.
	\item The hydrogen bonding network formed by the water may play a role 
        in structuring the pore.
	\item We can use simulated X-ray diffraction patterns to see if there is
        any meaningful structural difference between a "dry" and "wet" system
   \end{itemize}
    \end{enumerate}
%MRS2: size and how well well defined the pore is (how suddenly the RDF changes in the pore) is also a question.

%MRS2: hmm.  This last point still seems somewhat relevant.  Why left out?
%     \item While we might be able to provide answers to these questions using a 
%     molecular model, there remains the possibility that there is more than one 
%     metastable state associated with a given LLC system.

@@

  \begin{itemize}
    \item We validate the model using as much experimental information as possible.
    \item We are most interested in reproducing the conclusions about structure
    which have been made from X-ray diffraction (XRD) experiments and in matching ionic
    conductivity measurements~\cite{feng_thin_2016}.
    \begin{itemize}
	    \item We have compared simulated X-ray diffraction patterns to experiment in 
	    order to match major features present in the 2D patterns.
	    \item We calculated ionic conductivity using two agreeing methods.
	    \item We examined the the influence of crosslinking on membrane structure.
    \end{itemize}
    \item The structure-building approach and analysis used in this paper can be readily extended
    to the H\textsubscript{II} phase and other similar LC systems.
  \end{itemize}
%MRS2: and potentially used to help design membranes.  
  \section*{Methods}

  Liquid crystal monomers were parameterized using the Generalized AMBER Forcefield
  \cite{wang_development_2004} with the Antechamber package \cite{wang_automatic_2006}
  provided with AmberTools16 \cite{case_ambertools16_2016}. Atomic charges were
  assigned using tools from Openeye Scientific Software. 
%MRS2: be specific what tools
All molecular dynamics 
  simulations were run using the latest version of Gromacs 2016. %MRS2: give specific version, since latest may be different when this is published.
  \cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}

  An ensemble of characteristic, low-energy vacuum monomer configurations
  were constructed by applying a simulated annealing process to a
  parameterized monomer.
  \begin{itemize}
    \item Monomers were cooled from 1000K to 50K over 10 nanoseconds.
    \item A low energy configuration was randomly pulled from the trajectory
    and charges were reassigned using the am1bccsym method of molcharge
    shipped with Openeye Scientific software's QUACPAC % BJC: how to cite? %MRS2: webpage if no publication, 
    \item Using the new charges, the monomer system was annealed again and monomer
    configurations were pulled from the trajectory to be used for full
    system construction (Figure~\ref{fig:python}a).
  \end{itemize}

@@

  A typical simulation volume contains four pores in a monoclinic unit cell,
  the smallest unit cell that maintains hexagonal symmetry when extended 
  periodically.
  \begin{itemize}
    \item Each pore is made of twenty stacked monomer layers with periodic 
    continuity in the z direction, avoiding any edge effects and creating an 
    infinite length pore ideal for studying transport.
    \item A small number of layers is preferred in order to reduce computational
    cost and to allow us to look at longer timescales.
    \item Ultimately, we chose to build a system with 20 monomer layers in each pore
    in order to obtain sufficient resolution when simulating X-ray diffraction patterns.
    This point will be explained in more detail later.
    %MRS: you should give the reasons in the same place you describe the choice, but the supporting data would probably be better in supporting information
    %BJC: so refer to supporting info here? Then put plots of diffraction data with different sized system there. I did not explain the resolution point here because I haven't talked about simulated X-ray diffraction yet and it wouldn't make sense. And it doesn't make sense to talk about XRD this early in the methods but I need to justify the number of layers.
    %MRS2: I was referring to the data choice of 20 vs 10.  Yes, you can refer to it here, because the expectation is that peple will only look at it after reading the whole paper.
    \item We chose initial guesses for the remaining structural parameters based on
    experimental data and treated them as variables during model development.
  \end{itemize}

@@

%MRS2: make it clear these choices are for INITIAL spacing, and that this adjusted based on MD.  Also want to make
sure the dependence of the initial choices is described in supporting information, as the results shouldn't be TOO dependnt on thesechoices.
  We chose the initial layer spacing based on R-$\pi$.
  \begin{itemize}  
    \item Each monomer was rotated so the plane of the aromatic head groups would
    be parallel to the xy plane.
%MRS2: seems like the choice of separation distance here should be detailed in the supporting information, and how it changes  the structures formed.  Also, not quite clear. 
    \item The layers are placed so aromatic rings are stacked 3.7 \angstrom~
    apart in the z-direction.
    \item To quantify the degree of layering in our system, we calculated
    a spatial correlation function, $g(z)$, measured along the z-axis 
    (perpendicular to the membrane plane)
    \item To calculate $g(z)$, we binned the z component distances between 
    the center of mass of each benzene ring and all others of the same pore 
    over 50 ns of equilibrated trajectory and then normalized by the average
    number density.
    \item We compare the degree of layering between systems based on the
    amplitude of the first peak in $g(z)$. We halve the difference between the
    maximum of the first peak and the following minimum. We compare the difference
    to the mean to get a percentage deviation from the average number density.
    \item To extract the average distance between layers we applied a discrete 
    fourier transform to $g(z)$ and extracted the highest intensity frequency 
  \end{itemize}

  We placed pores at a chosen initial spacing based on R-pores, then allowed
  the system to settle into its preferred spacing. 
  \begin{itemize}
    %MRS: somewhere, should explain there's a large dependence of (meta)stable structures based on this initial distance.  It's sort of important to put somewhere.
    %BJC: Sure, but different choices of pore spacing were based on number of monomers per layer. More monomers per layer means a larger initial pore spacing.
    \item The model's pore centers are spaced 4.5 nm apart initially, $\approx$ 10 \% larger
    than the experimental value of 4.12 nm in order to reduce unintended repulsions 
    resulting from a tightly packed initial configuration.
    \item To calculate the equilibrated pore spacing, we measured the distance between pore
    centers.
    \item Pore centers were located by averaging the coordinates of sodium ions in their 
    respective pores.
    \item Statistics were generated using the bootstrapping technique (See Supplemental Information)
    %BJC: Do bootstrapping details go in supplemental?
    %MRS2: yes.
    	% Supplemental vvvv
	\begin{itemize}
        %MRS2: make clear sampling what? 
    	\item For each bootstrap trial, we recreate an equilibrium trajectory by randomly 
    	sampling from the original trajectory
    	\item Each pore spacing has its own trajectory with its own average value
    	\item The average value of each pore spacing is averaged to get the overall
    	average pore spacing. 
    	\item The standard deviation of average values is reported as the uncertainty in pore spacing
    	% BJC: The following will go in supplemental
        \item We are interested in 5 pore-to-pore distances which should all be equal in 
        a perfect hexagonal array. Only 4 are independent  
        % MRS: need to deal with the fact only 4 are independent in the bootstrapping?  How did we do that?
        % Supplemental ^^^^
	\end{itemize}
  \end{itemize}

@@

%MRS2: is the pore radius of the initial configuration really relevant? Does it need to be defined?  Seems like what radius means in the final equilibrated one is more important.
  \begin{figure}
 	\centering
 	\includegraphics[width=\textwidth]{pore_radius_illustration.png}
	\caption{The pore radius of the initial configuration is defined using the 
        carboxylate carbon as a reference}~\label{fig:pore_radius_illustration}
  \end{figure}

@@

  %We tested a number of equilibration schemes of increasing complexity in an effort
  %to match experiment while overcoming kinetic limitations that might trap the system 
  %in a metastable states. All start from an initial configuration generated with 
  %chosen structure variables.
  We developed equilibration schemes to create dry and wet configurations. Both schemes 
  start with an initial configuration generated according to the previous guidelines.
  \begin{itemize}
      \item Dry equilibration scheme:
      \begin{itemize}
        %MRS2: what initial force constant, and does it matter (that can be in supplemental) 
          \item Restraints fix monomer head groups in the sandwiched or parallel-displaced
          configurations while allowing monomer tails to settle.
          \item Doing so also mitigates system dependence on initial monomer configuration.
          \item The restrained portion of the equilibration scheme is run in the NVT ensemble.
          \item Every 50 ps, we reduce the force constants by the square root of its
          previous value, starting from 1e6 KJ mol$^-1$ nm$^-2$.
          \item Once the force constant is below 10 KJ mol$^-1$ nm$^-2$, the restraints are
          slowly released 
          %MRS2: be more quantitative about what ``slowly'' means.
          until there is no more restraining potential.
          \item The resulting unrestrained structure is allowed to equilibrate for 5 ns 
	  in the NPT ensemble with pressure controlled by the berendsen barostat.
	  \item Long, NPT equilibration simulations are run for at least 400 ns using the 
          Parrinello-Rahman barostat with a time consant of 10 ps.
      \end{itemize}
      \item Wet equilibration scheme: 
      \begin{itemize}
	  % BJC: the following three bullet points ended up not being the case since I used significantly less water
          %\item We obtained an estimate of equilibrium pore water content by solvating
          % an initial configuration with water baths above and below the membrane. 
	  %\item We ran the solvated system according to Scheme A but with a 1000 ns equilibration
          %\item Using that number we added water to the pores of an initial 
	  %configuration and equilibrated according to scheme B. 
	  \item In order to create a wet system, we solvated an initial configuration with water using gmx solvate
          \item All waters placed outside the pore region are removed 
	  \item Waters inside the pore region are randomly removed until the desired concentration
	  of water in the pores is reached. 
	  \item The remainder of the equilibration is the same as the dry system
      \end{itemize}
      \item In all cases, %MRS2: for system equilibration?
the v-rescale thermostat was used with tau-t = 0.1 % BJC: supplemental? %MRS2: probably here.
  \end{itemize}

>>>>

  While RO and NF have seen many advances in the past few decades, they 
  are far from perfect separation technologies.
  \begin{itemize}
    \item \textit{RO membranes}
    \item Inconsistent performance : Current state-of-the-art RO membranes are unstructured with
    tortuous and polydisperse diffusion pathways which leads to 
    inconsistent performance \cite{song_nano_2011}
    \item High energy requirements : Necessarily high feed pressures 
    drive up energy requirements which strains developing regions and
    contributes strongly to CO\textsubscript{2} emissions. \cite{mcginnis_global_2008}
    \item Separation based on differences in solubility and diffusivity:
%MRS2: couldn't you argue that the fact that they diffuse and dissolve at different rates means that it makes it possible to design targeted separations?
%BJC: right, I say something like that in the sentence following this one. It would be more challenging to design a targeted separation 
% especially if the feed is complex
    Moreover, designing RO membranes to achieve targeted separations of 
    specific solutes in a complex feed solution is nearly impossible because various solutes dissolve
    into and diffuse through the polymer matrix at different rates. \cite{wijmans_solution-diffusion_1995}
    \item With optimization, one can exploit these differences to create a functional
    selective barrier.
    \item \textit{NF membranes}
    \item NF was introduced as an intermediate between RO and ultrafiltration,
    having the ability to separate organic matter and salts on the order of 
    one nanometer in size.
%MRS2: is ``well-defined'' the right word if there is a ``pore size distribution''?
%BJC: By well-defined, I'm thinking of cylinders versus polymer matrix. One is definitely more
% defined than the other. What is the right word for that? unambiguous? explicit?
    \item Larger and well-defined pores drive down energy requirements while
    still affording separation of solutes as small as ions to some degree \cite{van_der_bruggen_review_2003}
    \item NF is often used as a precursor to reverse osmosis
    \item Unfortunately, NF membranes, like RO, possess a pore size 
    distribution which limits their ability to perform precise separations \cite{bowen_modelling_2002}
  \end{itemize}

  Nanostructured membranes can bypass many of the performance issues which
  plague traditional NF and RO membranes.
  \begin{itemize}
    \item Tune size and functionality of building blocks to control pore
    size and shape: One can accomplish targeted separations with high 
    selectivity by tuning shape, size and functionality of the molecular
    building blocks which form these materials. 
    \item As a result, solute rejecting pores can have their sizes tuned
    uniformly, resulting in strict size cut-offs.
    \item Entirely different mechanisms may govern transport in a given
    nanostructured material which can inspire novel separation techniques.
  \end{itemize}

@@

  Self assembling lyotropic liquid crystals (LLCs) are a suitable candidate for
  aqueous separation applications. 
  \begin{itemize}
    \item LLCs share the characteristic ability of nanostructured membrane
    materials to create highly ordered structures with the added benefits
    of low cost and synthetic techniques feasible for 
    large scale production \cite{feng_scalable_2014}
    \item Neat liquid crystal monomer forms the thermotropic, Col\textsubscript{h}
    phase. The presence of water results in the H\textsubscript{II} phase.
    \item In both cases, monomers assemble into mesophases made of hexagonally
    packed, uniform size, cylinders with hydrophilic groups oriented inward
    towards the pore center and hydrophobic groups facing outward.
    \item H\textsubscript{II} and Col\textsubscript{h} phase systems created by
    the monomer named Na-GA3C11 have been extensively studied experimentally \cite{smith_ordered_1997, %BJC: IUPAC chemical name here?
    zhou_supported_2005,resel_h2-phase_2000,feng_scalable_2014,feng_thin_2016}. 
  \end{itemize}

  Research into LLC membranes has been revived in recent years.
  \begin{itemize}
    \item During early stages of exploration, mesophases formed by Na-GA3C11 could not be macroscopically
    aligned, resulting in low flux membranes, and no clear route towards 
    scalable and economical filtration. 
    \item In 2014, Feng et al. showed that the mesophases could be aligned 
    using a magnetic field with subsequent crosslinking to lock the structure
    in place \cite{feng_scalable_2014}
    \item In 2016, Feng et al. showed that the same result could be obtained 
    using a technique termed soft confinement \cite{feng_thin_2016}.
  \end{itemize}

@@

  A molecular level understanding of LLC membrane structure, enabled by molecular
  dynamics simulations, will provide guidelines to reduce the large chemical space
  available to design monomers for creation of separation-specific membranes.
  \begin{itemize}
    \item Over the past 20 years, H\textsubscript{II} phase LLC membrane studies have been limited primarily
    to Na-GA3C11 with some characterization done after minor structural modifications
    \item Resel et al. varied the length of the monomer tails and the counterion used and 
    observed its affect on pore spacing \cite{resel_structural_2000}.
    \item Rejection studies show that this membrane can not perform separations of
    solutes less than 1.2 nm in diameter because the pores are too large \cite{zhou_supported_2005}.
    \item  We do not yet understand how to controllably reduce the effective pore size or 
    how to tune the chemical environment in the nanopores for effective water
    desalination or small organic molecule separations.
    \item It will be challenging to efficiently narrow down the large design space in 
    a laboratory setting without a robust model.
    \item The only source of predictive modeling for LLC systems have been macroscopic models 
    which likely do not adequately describe transport at these length scales. % BJC: Reference. I think it is just w.r.t. bicontinuous cubic
%MRS2: not quite sure of the phrasing here on next two points. Something more a long the lines of the choice of head group 
%may allow tuning rejection of solutes based on size?
%BJC: okay tried again
    \item The choice of head group may allow us to tune pore size for size exclusion driven separations
    \item Choice of counterion may influence the establishment of a Donnan potential
    affecting the degree to which the membrane can exclude charged species.
    \item A good molecular model should incorporate a detailed picture of the nanoscopic pore 
    structure which will be crucial to understanding the role of monomer structure in 
    membrane design.
    \item Molecular dynamics simulations will provide the required level of detail
  \end{itemize}

  %BJC: Figure here showing previous level of detail compared to atomistic level of detail
  %MRS2: provenance of the atomistic figure could be explained better in the caption. It also looks disordered. Will eventually need a higher resolution figure for the previous understanding pore.
  %BJC: Okay. The picture here is of the "ordered pore" phase. It looks disorderd from this angle. I could use a different slice of the membrane to see if it looks any better
  \begin{figure}
  \centering
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{nanopore_undetailed.jpg}
		\caption{}~\label{fig:undetailed_pore}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{detailed_pore.png}
		\caption{}~\label{fig:detailed_pore}
	\end{subfigure}
  \caption{(a) Our previous understanding of the pore structure allows us to speculate
	   about separation behavior. (b) A detailed molecular model will allow us to
	   directly observe solute transport. Here, four stacked layers of 5 monomers
           are pictured atomistically. The hydrophilic region is in red and the 
	   hydrophoic region is colored blue.}~\label{fig:detail}
  \end{figure}

  Our approach to constructing a general model will follow the development of a
  model of a specific LLC membrane with sufficient experimental characterization.
    \begin{itemize}
	    \item We have chosen to focus on assemblies formed by Na-GA3C11
	    \item We have also narrowed our scope to the development of 
	    a model of the Col\textsubscript{h} phase membrane.
	    \item Compared to the H\textsubscript{II} phase, the Col\textsubscript{h}
	    phase is a simpler starting point, due to the absence of water, and has
	    detailed experimental WAXS patterns useful for reconstructing structural data.  
    \end{itemize}

@@

   \item Is it necessary to include any water in order to appropriately model the 
   Col\textsubscript{h} phase? \label{point:water}
   \begin{itemize}
	\item While the Col\textsubscript{h} phase is described as dry, it is 
	possible that small amounts of ambient water may be leached into the system.
	\item The hydrogen bonding network formed by the water may play a role 
        in structuring the pore.
	\item We can use simulated X-ray diffraction patterns to see if there is
        any meaningful structural difference between a "dry" and "wet" system
   \end{itemize}
    \end{enumerate}
%MRS2: size and how well defined the pore is (how suddenly the RDF changes in the pore) is also a question.

%MRS2: hmm.  This last point still seems somewhat relevant.  Why left out?
%BJC: I thought we agreed to leave this out. It lengthens the paper and makes it
% more confusing. Unless we just want to hint at it.

%     \item While we might be able to provide answers to these questions using a 
%     molecular model, there remains the possibility that there is more than one 
%     metastable state associated with a given LLC system.

@@

  \begin{itemize}
    \item We validate the model using as much experimental information as possible.
    \item We are most interested in reproducing the conclusions about structure
    which have been made from X-ray diffraction (XRD) experiments and in matching ionic
    conductivity measurements~\cite{feng_thin_2016}.
    \begin{itemize}
	    \item We have compared simulated X-ray diffraction patterns to experiment in 
	    order to match major features present in the 2D patterns.
	    \item We calculated ionic conductivity using two agreeing methods.
	    \item We examined the the influence of crosslinking on membrane structure.
    \end{itemize}
    \item The structure-building approach and analysis used in this paper can be readily extended
    to the H\textsubscript{II} phase, other similar LC systems and ultimately to help design
    real membranes.
  \end{itemize}

  \section*{Methods}

  Liquid crystal monomers were parameterized using the Generalized AMBER Forcefield
  \cite{wang_development_2004} with the Antechamber package \cite{wang_automatic_2006}
  provided with AmberTools16 \cite{case_ambertools16_2016}. Atomic charges were
  assigned using the am1bccsym method of molcharge shipped with QUACPAC from Openeye % cite with webpage
  Scientific Software. All molecular dynamics simulations were run using Gromacs 2016  % This the version on Bridges
  \cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}

  An ensemble of characteristic, low-energy vacuum monomer configurations
  were constructed by applying a simulated annealing process to a
  parameterized monomer.
  \begin{itemize}
    \item Monomers were cooled from 1000K to 50K over 10 nanoseconds.
    \item A low energy configuration was randomly pulled from the trajectory
    and charges were reassigned with molcharge. 
    \item Using the new charges, the monomer system was annealed again and monomer
    configurations were pulled from the trajectory to be used for full
    system construction (Figure~\ref{fig:python}a).
  \end{itemize}

@@

  A typical simulation volume contains four pores in a monoclinic unit cell,
  the smallest unit cell that maintains hexagonal symmetry when extended 
  periodically.
  \begin{itemize}
    \item Each pore is made of twenty stacked monomer layers with periodic 
    continuity in the z direction, avoiding any edge effects and creating an 
    infinite length pore ideal for studying transport.
    \item Each layer is made of a set number of monomers per layer, a variable to
    be explored further.
    \item A small number of layers is preferred in order to reduce computational
    cost and to allow us to look at longer timescales.
    \item Ultimately, we chose to build a system with 20 monomer layers in each pore
    in order to obtain sufficient resolution when simulating X-ray diffraction patterns.
    This point will be explained in more detail later.
    %MRS: you should give the reasons in the same place you describe the choice, but the supporting data would probably be better in supporting information
    %BJC: so refer to supporting info here? Then put plots of diffraction data with different sized system there. I did not explain the resolution point here because I haven't talked about simulated X-ray diffraction yet and it wouldn't make sense. And it doesn't make sense to talk about XRD this early in the methods but I need to justify the number of layers.
    %MRS2: I was referring to the data choice of 20 vs 10.  Yes, you can refer to it here, because the expectation is that peple will only look at it after reading the whole paper.
    \item We chose initial guesses for the remaining structural parameters based on
    experimental data and treated them as variables during model development.
  \end{itemize}

@@

%MRS2: make it clear these choices are for INITIAL spacing, and that this adjusted based on MD.  Also want to make
%sure the dependence of the initial choices is described in supporting information, as the results shouldn't be TOO dependnt on thesechoices.
%BJC: okay will work on some figures illustrating the affect of initial choices on results
  We chose the layer spacing for the initial configuration based on R-$\pi$ and then allowed system
  to readjust during equilibration.
  \begin{itemize}  
    \item Each monomer was rotated so the plane of the aromatic head groups would
    be parallel to the xy plane.
%MRS2: seems like the choice of separation distance here should be detailed in the supporting information, and how it changes  the structures formed.  Also, not quite clear. 
%BJC: okay, I have some ideas for this
    \item The layers are placed so aromatic rings are stacked 3.7 \angstrom~
    apart in the z-direction.
    \item To quantify the degree of layering in our system, we calculated
    a spatial correlation function, $g(z)$, measured along the z-axis 
    (perpendicular to the membrane plane)
    \item To calculate $g(z)$, we binned the z component distances between 
    the center of mass of each benzene ring and all others of the same pore 
    over 50 ns of equilibrated trajectory and then normalized by the average
    number density.
    \item We compare the degree of layering between systems based on the
    amplitude of the first peak in $g(z)$. We halve the difference between the
    maximum of the first peak and the following minimum. We compare the difference
    to the mean to get a percentage deviation from the average number density.
    \item To extract the average distance between layers we applied a discrete 
    fourier transform to $g(z)$ and extracted the highest intensity frequency 
  \end{itemize}

  We placed pores at a chosen initial spacing based on R-pores, then allowed
  the system to settle into its preferred spacing. 
  \begin{itemize}
    %MRS: somewhere, should explain there's a large dependence of (meta)stable structures based on this initial distance.  It's sort of important to put somewhere.
    %BJC: Sure, but different choices of pore spacing were based on number of monomers per layer. More monomers per layer means a larger initial pore spacing.
    \item The model's pore centers are spaced 4.5 nm apart initially, $\approx$ 10 \% larger
    than the experimental value of 4.12 nm in order to reduce unintended repulsions 
    resulting from a tightly packed initial configuration.
    \item To calculate the equilibrated pore spacing, we measured the distance between pore
    centers.
    \item Pore centers were located by averaging the coordinates of sodium ions in their 
    respective pores.
    \item Statistics were generated using the bootstrapping technique (See Supplemental Information)
    	% Supplemental vvvv
	\begin{itemize}
        %MRS2: make clear sampling what?
        \item We are interested in 5 pore-to-pore distances which should all be equal in
        a perfect hexagonal array, however only 4 distances are independent % can visualize this in supplemental material. Will make a lot of things more clear below.
	\item Each pore spacing has its own trajectory of spacing vs. time. Using data collected after
	the system is equilibrated, we calculate how long it takes for the data in each of the 5
	trajetories to become uncorrelated using pymbar.timeseries.integratedAutocorrelationTime() % reference
	\item We break the full trajectories down into sub-trajectories based on the
	maximum autocorrelation time of those found in the previous step. 
    	\item For each bootstrap trial, we recreate an equilibrium trajectory by randomly 
    	sampling pore spacings from the sub-trajectories 
    	\item We get an average value for each pore spacing by finding the mean of the bootstrapped data
	\item We calculate the overall average as the mean of all bootstrapped pore spacings
	\item The uncertainty for each pore spacing is calculated as $\dfrac{<x> - x}{4}$ where $<x>$ is the
	average spacing from the bootstrap trial and $x$ is the average value of one of the pore spacings.
	\item We report the mean of these uncertainties
        % MRS: need to deal with the fact only 4 are independent in the bootstrapping?  How did we do that?
	% BJC: Updated
        % Supplemental ^^^^
	\end{itemize}
  \end{itemize}

@@

%MRS2: is the pore radius of the initial configuration really relevant? Does it need to be defined?  Seems like what radius means in the final equilibrated one is more important.
%BJC: maybe not, we'll have to talk to get on the same page
  \begin{figure}
 	\centering
 	\includegraphics[width=\textwidth]{pore_radius_illustration.png}
	\caption{The pore radius of the initial configuration is defined using the 
        carboxylate carbon as a reference}~\label{fig:pore_radius_illustration}
  \end{figure}

@@

  %We tested a number of equilibration schemes of increasing complexity in an effort
  %to match experiment while overcoming kinetic limitations that might trap the system 
  %in a metastable states. All start from an initial configuration generated with 
  %chosen structure variables.
  We developed equilibration schemes to create dry and wet configurations. Both schemes 
  start with an initial configuration generated according to the previous guidelines.
  \begin{itemize}
      \item Dry equilibration scheme:
      \begin{itemize}
        %MRS2: what initial force constant, and does it matter (that can be in supplemental) 
	%BJC: I've never explored whether there is any dependence on initial force constant. It probably doesn't matter unless it starts really low
          \item Restraints fix monomer head groups in the sandwiched or parallel-displaced
          configurations while allowing monomer tails to settle.
          \item Doing so also mitigates system dependence on initial monomer configuration.
          \item The restrained portion of the equilibration scheme is run in the NVT ensemble.
          \item Every 50 ps, we reduce the force constants by the square root of its
          previous value, starting from 1e6 KJ mol$^-1$ nm$^-2$.
          \item Once the force constant is below 10 KJ mol$^-1$ nm$^-2$, the restraints are
          released linearly until there is no more restraining potential.
          \item The resulting unrestrained structure is allowed to equilibrate for 5 ns 
	  in the NPT ensemble with pressure controlled by the berendsen barostat.
	  \item Long, NPT equilibration simulations are run for at least 400 ns using the 
          Parrinello-Rahman barostat with a time consant of 10 ps.
      \end{itemize}
      \item Wet equilibration scheme: 
      \begin{itemize}
	  % BJC: the following three bullet points ended up not being the case since I used significantly less water
          %\item We obtained an estimate of equilibrium pore water content by solvating
          % an initial configuration with water baths above and below the membrane. 
	  %\item We ran the solvated system according to Scheme A but with a 1000 ns equilibration
          %\item Using that number we added water to the pores of an initial 
	  %configuration and equilibrated according to scheme B. 
	  \item In order to create a wet system, we solvated an initial configuration with water using gmx solvate
          \item All waters placed outside the pore region are removed 
	  \item Waters inside the pore region are randomly removed until the desired concentration
	  of water in the pores is reached. 
	  \item The remainder of the equilibration is the same as the dry system
      \end{itemize}
      \item In all cases, the v-rescale thermostat was used with tau-t = 0.1 % BJC: supplemental
  \end{itemize}
